---
title: "ca"
output: html_document
date: "2026-01-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 11. Then, create an RMarkdown file that knits to an HTML file which

Ah look, here we are! ...gonna install/load dependencies here.

```{r}
# just going to install all of tidyverse
# install.packages("tidyverse")
library(tidyverse)
```

## 12. Reads the two allele frequency files and the logistic regression output into R, and

```{r}
lr <- read.table(
  "data/data.assoc.logistic",
  header = TRUE
)
ctrls <- read.table(
  "data/data4.ctrls.freq",
  header = TRUE
)
cases <- read.table(
  "data/data4.cases.freq",
  header = TRUE
)
```

## 13. Plots a histogram of the case allele frequencies,

```{r}
cases %>% 
  ggplot() +
  aes(x = MAF) +
  geom_histogram(bins = 20) +
  theme_minimal()
```

## 14. Plots a histogram of the control allele frequencies,

```{r}
ctrls %>% 
  ggplot() +
  aes(x = MAF) +
  geom_histogram(bins = 20) +
  theme_minimal()
```

## 15. Plots a scatterplot of the case individuals’ allele frequencies (y axis) --AND-- 16. Versus the control individuals’ allele frequencies (x axis),

```{r}
# merge the cases and controls on the SNP
case_controls <- cases %>% 
  select(SNP, MAF) %>% 
  rename(`Case MAF` = MAF) %>%
  inner_join(
    ctrls %>%
      select(SNP, MAF) %>%
      rename(`Control MAF` = MAF),
    by = "SNP"
  )
# create the scatterplot
case_controls %>% 
  ggplot() +
  aes(x = `Control MAF`, y = `Case MAF`) +
  geom_point()
```

## 17. Plots a scatterplot of −log₁₀(p values) (y axis) versus chr4 basepair position (x axis), --AND-- 18. Reports the marker and p value of the marker with the lowest p value; then

```{r}
lr$NegLogP <- -log10(lr$P)
lr_cleaned <- lr %>% 
  drop_na() %>% 
  filter(CHR == 4) %>% 
  mutate(color = factor(if_else(P == min(P), "Min", "Other")))
lr_cleaned %>% 
  ggplot() +
  aes(
    y = NegLogP,
    x = BP,
    color = color
  ) +
  geom_point() +
  # source referenced for annotating the min point:
  # https://stackoverflow.com/questions/14351608/color-one-point-and-add-an-annotation-in-ggplot2
  scale_color_manual(values = c("Min" = "red", "Other" = "black")) +
  geom_text(data = textdf, aes(x = NegLogP * 1.05, y = BP, label = "my label"))
  
```

## 19. Loads the snp_annot.RData workspace and merges the p values into the SNP annotation data frame object snp_annot_df in such a way that --AND-- 20. The order of the SNPs in snp_annot_df is unchanged.

```{r}

```
